Structured Versus Decorated Cospans

### 2020-12-08 cospans.tex
various edits and comments up to 5.4

	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=inarrow, bend left=20, looseness=1.00] (n1) to (n2);
		\draw [style=inarrow, bend right=20, looseness=1.00] (n1) to (n3);
		\draw [style=inarrow, bend left=20, looseness=1.00] (n2) to (n4);
		\draw [style=inarrow, bend right=20, looseness=1.00] (n3) to (n4);
		\draw [style=inarrow] (n2) to (n3);
%		\draw [style=inarrow] (W) to (Water);
%		\draw [style=inarrow, bend left=40, looseness=1.00] (Water2) to (Something);
%		\draw [style=inarrow, bend right=40, looseness=1.00] (Water2) to (Something);
%		\draw [style=inarrow, bend left=40, looseness=1.00] (Something) to (A);
%		\draw [style=inarrow, bend right=40, looseness=1.00] (Something) to (B);
		\draw [style=simple] (ATL.center) to (ATR.center);
		\draw [style=simple] (ATR.center) to (ABR.center);
		\draw [style=simple] (ABR.center) to (ABL.center);
		\draw [style=simple] (ABL.center) to (ATL.center);
%		\draw [style=simple] (BTL.center) to (BTR.center);
%		\draw [style=simple] (BTR.center) to (BBR.center);
%		\draw [style=simple] (BBR.center) to (BBL.center);
%		\draw [style=simple] (BBL.center) to (BTL.center);
		\draw [style=simple] (MTL.center) to (MTR.center);
		\draw [style=simple] (MTR.center) to (MBR.center);
		\draw [style=simple] (MBR.center) to (MBL.center);
		\draw [style=simple] (MBL.center) to (MTL.center);
%		\draw [style=inputarrow] (outI) to (A);
%		\draw [style=inputarrow] (outS) to (B);
		\draw [style=inputarrow] (inI) to (n1);
		\draw [style=inputarrow] (outI') to (n4);
%		\draw [style=inputarrow] (inI') to (Water2);
%		\draw [style=inputarrow] (inS') to (Water2);
	\end{pgfonlayer}
\end{tikzpicture}
}
\]
We might hope to do this using a symmetric lax monoidal functor $F \maps (\Fin\Set, +) \to (\Set, \times)$ assigning to each finite set $m$ the set of all graphs with $m$ as their set of nodes. But this hope is doomed, for reasons painstakingly explained in \cite[Sec.\ 5]{BC2}. 
There is really a \emph{category} of graphs with $m$ as their set of nodes---and surprisingly, trying to treat this category as a mere set does not work, despite all the tricks one might try.

@@

In fact another solution to the problem is already known: the theory of structured cospans \cite{BC2,CourserThesis}.  Given a functor $L \maps \A \to \X$, a \define{structured cospan} is a cospan in $\X$ whose feet come from a pair of objects in $\A$:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,1) {$x$};
\node (C) at (2,0) {$L(b).$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C)edge node[above]{$$}(B);
\end{tikzpicture}
\]
This is another way of letting the apex have more structure than the feet.   When $\A$ and $\X$ have finite colimits and $L$ preserves them, there is a symmetric monoidal double category ${}_L \lCsp(\X)$ where:
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a diagram in $\X$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C)edge node[above]{$o$}(B);
\end{tikzpicture}
\]
\item a 2-morphism is a commutative diagram in $\X$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,0) {$L(a)$};
\node (F) at (5,0) {$L(b)$};
\node (G) at (4,0) {$x$};
\node (E') at (3,-1) {$L(a')$};
\node (F') at (5,-1) {$L(b')$};
\node (G') at (4,-1) {$x'$};
\path[->,font=\scriptsize,>=angle 90]
(F) edge node[above]{$o$} (G)
(E) edge node[left]{$L(f)$} (E')
(F) edge node[right]{$L(g)$} (F')
(G) edge node[left]{$\alpha$} (G')
(E) edge node[above]{$i$} (G)
(E') edge node[below]{$i'$} (G')
(F') edge node[below]{$o'$} (G');
\end{tikzpicture}
\]
\end{itemize}
Many of the flawed applications of decorated cospans have been fixed using structured cospans \cite[Sec.\ 6]{BC2}, but not all decorated cospans can handled using structured cospans.

@@

The first part formulates the existence of colimits \emph{locally} in each fiber, and if we let $\J$ range over all finite categories it says that the corresponding pseudofunctor $F\maps\A\to\Cat$ lands in the sub-2-category $\Rex$. The second part formulates the existence of colimits \emph{globally} in the total category $\inta F$, and if we let $\J$ range over all finite categories it says that $\X$ has finite colimits and $U$ preserves all finite colimits. As an example, suppose that $\A$ has pushouts and a Grothendieck opfibration $U\colon\inta F\to\A$ has fibrewise pushouts preserved by its reindexing functors. We can construct pushouts in $\inta F$ as follows: 
\begin{equation}\label{eq:globalpushout}
 \begin{tikzcd}[ampersand replacement=\&,column sep=1in,row sep=.5in]
\& (a+_bc,w)\ar[dd,phantom, very near start,"\upback"description] \& \\
(a,x)\ar[ur] \&\& (c,z)\ar[ul] \\
\& (b,y)\ar[ul,"{\begin{cases}f\maps b\to a &\textrm{in }\A \\k\maps F(f)y\to x &\textrm{in }F(a)\end{cases}}"description]\ar[ur,"{\begin{cases}g\maps b\to c &\textrm{in }\A \\\ell\maps F(g)y\to z &\textrm{in }F(c)\end{cases}}"description] \&
 \end{tikzcd}
\end{equation}
which on the first variable is given by the pushout in $\A$ on the left below, and the object $w$ is formed as the pushout in the fiber category $F(a+_bc)$ on the right below:
\begin{displaymath}
\begin{tikzcd}
& a+_bc\ar[dd,phantom, very near start,"\upback"description] & \\
a\ar[ur,"u_a"] && c\ar[ul,"u_c"'] \\
& b\ar[ul,"f"]\ar[ur,"g"'] &
\end{tikzcd}\qquad
\begin{tikzcd}
 & w \ar[dd,phantom, very near start,"\upback"description] & \\
 F(u_a)x\ar[ur] & & F(u_c)z\ar[ul] \\
 & F(u_a)(Ff(y))\cong F(u_c)(Fg(y))\ar[ul,"F(u_a)k"]\ar[ur,"F(u_c)\ell"'] &
 \end{tikzcd}
\end{displaymath}
In a similar way we can construct the initial object and coproducts in a Grothendieck category from those in the fibers, namely
\begin{equation}\label{eq:fibrewisecoprod}
(a,x)+(b,y)=\left(a+ b,F(\iota_a)x+F(\iota_b)y\right), \qquad 
 0_{\inta F}=\left(0_A,0_{F(0_A)}\right)
\end{equation}
where $\iota_a:a\to a+b\leftarrow b:\iota_b$ are the sum inclusions in $\A$ and the sum on the second variable is in the fiber $F(a+b)$.

@@

Notice a subtlety in the above formulation. \cref{lem:MonGroth} provides a specific monoidal structure on the fibers which becomes \cref{eq:explicitstructure1} when the base $\A$ is cocartesian monoidal.   If this fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to any other cocartesian monoidal structure on the fibers, by the universal property of colimits. Therefore, we can simplify the above assumption to that in \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor, although we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$. For example, in that case, the coproducts induced in the Grothendieck category from the fibrewise ones as in \cref{eq:fibrewisecoprod} are equivalently given by \cref{eq:explicitstructure2}.

@@

Building up to an isomorphism of the above described double categories, we now define a double functor
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp
\end{displaymath}
which on vertical categories acts as the identity $\lE_0 = \id_{\A}$. Given a horizontal 1-cell in $_L\lCsp(\inta F)$ as in \cref{eq:scsphor1cell}, $\lE_1$ maps it to the decorated cospan $i\maps a\to m\leftarrow b:o$ with decoration $x\in F(m)$,
whereas a 2-morphism of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} is mapped to the underlying map of cospans in $\A$
\begin{displaymath}
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\end{displaymath}
in $\A$ along with the map $\tau\maps Fk(x)\to x'$ in $F(m')$ coming from the middle arrow above, namely an $F$-decorated cospan 2-morphism. 
With these definitions, $\lE$ is indeed a double functor as in \cref{def:doublefun}: both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets, and there exist globular 2-isomorphisms $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(N)\odot \mathbb{E}(M)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b,\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn)
\end{equation}
whereas first composing $M$ and $N$ in $_L\lCsp(\inta F)$ using pushouts in $\inta F$ as in \cref{eq:globalpushout} produces 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(u_m)x+_{(\bot_{m+_bn})}F(u_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $u$ are the canonical maps into the pushout.
Under $\lE$ this is mapped to the underlying $\A$-cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(u_m)x+F(u_n)y$, since a pushout over an initial object is really a coproduct. That decoration is isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}
F(m)\times F(n)\ar[rr,"\phi"]\ar[d,"F(u_m)\times F(u_n)"']\ar[dr,phantom,"\cong"] && F(m+n)\ar[d,"F(\psi)"]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(u_m+u_n)"description] \\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the bottom composite is the fibrewise coproduct in $F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand side isomorphism is due to pseudonaturality of $\phi$ whereas the right-hand side isomorphism follows from universal properties and pseudofunctoriality of $F$.
Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibers. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}. %Do we want to check that E is a double functor one last time?

@@

\begin{proof}
This follows by applying Shulman's result \cite[Thm.\ 1.2]{Shulman2010} to the fibrant symmetric monoidal double category $F\lCsp$.
\end{proof}

@@

Decorated cospans were introduced by Fong \cite{Fong,FongThesis} to describe open systems as cospans equipped with extra data.  They were then applied to open electrical circuits \cite{BF}, Markov processes \cite{BFP}, and chemical reaction networks \cite{BP}.  Unfortunately, some of these applications were marred by technical flaws, which were later fixed using structured cospans \cite[Sec.\ 5]{BC}.  Our new approach to decorated cospans provides another solution to these problems.  Here we compare the two approaches in applications to graphs, electrical circuits, Petri nets, reaction networks and dynamical systems.   In many cases \cref{thm:equiv} shows that the structured and decorated cospan approaches are equivalent, but we show that open dynamical systems can only be treated using decorated cospans.

@@

We can easily build a double category with `open graphs' as horizontal 1-cells using the machinery of structured cospans.  Let $L \maps \Fin\Set \to \Graph$ be the functor that assigns to a finite set $N$ the \define{discrete graph} on $N$: the graph with no edges and $N$ as its set of vertices. Both $\Fin\Set$ and $\Graph$ have finite colimits, and the functor $L \maps \Fin\Set \to \Graph$ is left adjoint to the forgetful functor $R \maps \Graph \to \Fin\Set$ that assigns to a graph $G$ its underlying set of vertices $R(G)$. Thus, using structured cospans and appealing to \cref{thm:SC}, we get a symmetric monoidal double category $_L \lCsp(\Graph)$ in which:
\begin{itemize}
\item objects are finite sets,
\item a vertical 1-morphism from $X$ to $Y$ is a function $f \maps X \to Y$,
\item a horizontal 1-cell from $X$ to $Y$ is an \define{open graph} from $X$ to $Y$, meaning a cospan in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a commuting diagram in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(Y_1)$};
\node (A') at (0,-1) {$L(X_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(Y_2)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
\end{itemize}
Here is an example of an open graph:
\[
\scalebox{0.8}{
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [contact] (n1) at (-2,0) {$\bullet$};
		\node [style = none] at (-2.1,0.3) {$n_1$};
		\node [contact] (n2) at (0,1) {$\bullet$};
		\node [style = none] at (0,1.3) {$n_2$};
		\node [contact] (n3) at (0,-1) {$\bullet$};
		\node [style = none] at (0,-1.3) {$n_3$};
		\node [contact] (n4) at (2,0) {$\bullet$};
		\node [style = none] at (2.1,0.3) {$n_4$};

@@

We can make $F$ into a symmetric lax monoidal pseudofunctor $F \maps (\Fin\Set, +) \to (\Cat,\times)$ by equipping it with suitable functors 
\[   \phi_{N,N'} \maps F(N) \times F(N') \to F(N+N'), \qquad \phi_0 \maps F(\emptyset) \to 1 \in \Cat. \]  
The functor $\phi_0$ is uniquely determined since $1$ is the terminal category.   More interesting is $\phi_{N,N'}$.   This functor maps a pair of graph structures $s, t \maps E \to N$ and $s',t' \maps E' \to N'$ to the graph structure $s+s', t+t' \maps E+E' \to N+N'$.  In other words, it sends a pair of graph structures to their `disjoint union'.   Surprisingly, though $F$ is a functor---for example, composition is strictly preserved---this choice of $\phi_{N,N'}$ does not make $F$ into a symmetric lax monoidal functor, since it obeys the required laws only up to natural isomorphism, as in \cref{eq:omega}.   See \cite[Section 5]{BC} for a proof that these laws fail to hold on the nose.   This fact is what necessitated a generalization of Fong's original approach to decorated cospans.

>>>>

	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=inarrow, bend left=20, looseness=1.00] (n1) to (n2);
		\draw [style=inarrow, bend right=20, looseness=1.00] (n1) to (n3);
		\draw [style=inarrow, bend left=20, looseness=1.00] (n2) to (n4);
		\draw [style=inarrow, bend right=20, looseness=1.00] (n3) to (n4);
		\draw [style=inarrow] (n2) to (n3);
%		\draw [style=inarrow] (W) to (Water);
%		\draw [style=inarrow, bend left=40, looseness=1.00] (Water2) to (Something);
%		\draw [style=inarrow, bend right=40, looseness=1.00] (Water2) to (Something);
%		\draw [style=inarrow, bend left=40, looseness=1.00] (Something) to (A);
%		\draw [style=inarrow, bend right=40, looseness=1.00] (Something) to (B);
		\draw [style=simple] (ATL.center) to (ATR.center);
		\draw [style=simple] (ATR.center) to (ABR.center);
		\draw [style=simple] (ABR.center) to (ABL.center);
		\draw [style=simple] (ABL.center) to (ATL.center);
%		\draw [style=simple] (BTL.center) to (BTR.center);
%		\draw [style=simple] (BTR.center) to (BBR.center);
%		\draw [style=simple] (BBR.center) to (BBL.center);
%		\draw [style=simple] (BBL.center) to (BTL.center);
		\draw [style=simple] (MTL.center) to (MTR.center);
		\draw [style=simple] (MTR.center) to (MBR.center);
		\draw [style=simple] (MBR.center) to (MBL.center);
		\draw [style=simple] (MBL.center) to (MTL.center);
%		\draw [style=inputarrow] (outI) to (A);
%		\draw [style=inputarrow] (outS) to (B);
		\draw [style=inputarrow] (inI) to (n1);
		\draw [style=inputarrow] (outI') to (n4);
%		\draw [style=inputarrow] (inI') to (Water2);
%		\draw [style=inputarrow] (inS') to (Water2);
	\end{pgfonlayer}
\end{tikzpicture}
}
\]
We might hope to do this using a symmetric lax monoidal functor $F \maps (\Fin\Set, +) \to (\Set, \times)$ assigning to each finite set $m$ the set of all graphs with $m$ as their set of nodes. But this hope is doomed, for reasons painstakingly explained in \cite[Section 5]{BC2}. 
There is really a \emph{category} of graphs with $m$ as their set of nodes---and surprisingly, trying to treat this category as a mere set does not work, despite all the tricks one might try.

@@

In fact another solution to the problem is already known: the theory of structured cospans \cite{BC2,CourserThesis}.  Given a functor $L \maps \A \to \X$, a \define{structured cospan} is a cospan in $\X$ whose feet come from a pair of objects in $\A$:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,1) {$x$};
\node (C) at (2,0) {$L(b).$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C)edge node[above]{$$}(B);
\end{tikzpicture}
\]
This is another way of letting the apex have more structure than the feet.   When $\A$ and $\X$ have finite colimits and $L$ preserves them, there is a symmetric monoidal double category ${}_L \lCsp(\X)$ where:
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a diagram in $\X$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C)edge node[above]{$o$}(B);
\end{tikzpicture}
\]
\item a 2-morphism is a commutative diagram in $\X$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,0) {$L(a)$};
\node (F) at (5,0) {$L(b)$};
\node (G) at (4,0) {$x$};
\node (E') at (3,-1) {$L(a')$};
\node (F') at (5,-1) {$L(b')$};
\node (G') at (4,-1) {$x'$};
\path[->,font=\scriptsize,>=angle 90]
(F) edge node[above]{$o$} (G)
(E) edge node[left]{$L(f)$} (E')
(F) edge node[right]{$L(g)$} (F')
(G) edge node[left]{$\alpha$} (G')
(E) edge node[above]{$i$} (G)
(E') edge node[below]{$i'$} (G')
(F') edge node[below]{$o'$} (G');
\end{tikzpicture}
\]
\end{itemize}
Many of the flawed applications of decorated cospans have been fixed using structured cospans \cite[Section 6]{BC2}, but not all decorated cospans can handled using structured cospans.

@@

The first part formulates the existence of colimits \emph{locally} in each fiber, and if we let $\J$ range over all finite categories it says that the corresponding pseudofunctor $F\maps\A\to\Cat$ lands in the sub-2-category $\Rex$. The second part formulates the existence of colimits \emph{globally} in the total category $\inta F$, and if we let $\J$ range over all finite categories it says that $\X$ has finite colimits and $U$ preserves all finite colimits. As an example, suppose that $\A$ has pushouts and a Grothendieck opfibration $U\colon\inta F\to\A$ has fiberwise pushouts preserved by its reindexing functors. We can construct pushouts in $\inta F$ as follows: 
\begin{equation}\label{eq:globalpushout}
 \begin{tikzcd}[ampersand replacement=\&,column sep=1in,row sep=.5in]
\& (a+_bc,w)\ar[dd,phantom, very near start,"\upback"description] \& \\
(a,x)\ar[ur] \&\& (c,z)\ar[ul] \\
\& (b,y)\ar[ul,"{\begin{cases}f\maps b\to a &\textrm{in }\A \\k\maps F(f)y\to x &\textrm{in }F(a)\end{cases}}"description]\ar[ur,"{\begin{cases}g\maps b\to c &\textrm{in }\A \\\ell\maps F(g)y\to z &\textrm{in }F(c)\end{cases}}"description] \&
 \end{tikzcd}
\end{equation}
which on the first variable is given by the pushout in $\A$ on the left below, and the object $w$ is formed as the pushout in the fiber category $F(a+_bc)$ on the right below:
\begin{displaymath}
\begin{tikzcd}
& a+_bc\ar[dd,phantom, very near start,"\upback"description] & \\
a\ar[ur,"u_a"] && c\ar[ul,"u_c"'] \\
& b\ar[ul,"f"]\ar[ur,"g"'] &
\end{tikzcd}\qquad
\begin{tikzcd}
 & w \ar[dd,phantom, very near start,"\upback"description] & \\
 F(u_a)x\ar[ur] & & F(u_c)z\ar[ul] \\
 & F(u_a)(Ff(y))\cong F(u_c)(Fg(y))\ar[ul,"F(u_a)k"]\ar[ur,"F(u_c)\ell"'] &
 \end{tikzcd}
\end{displaymath}
In a similar way we can construct the initial object and coproducts in a Grothendieck category from those in the fibers, namely
\begin{equation}\label{eq:fibrewisecoprod}
(a,x)+(b,y)=\left(a+ b,F(\iota_a)x+F(\iota_b)y\right), \qquad 
 0_{\inta F}=\left(0_A,0_{F(0_A)}\right)
\end{equation}
where $\iota_a:a\to a+b\leftarrow b:\iota_b$ are the sum inclusions in $\A$ and the sum on the second variable is in the fiber $F(a+b)$.

@@

Notice a subtlety in the above formulation. \cref{lem:MonGroth} provides a specific monoidal structure on the fibers which becomes \cref{eq:explicitstructure1} when the base $\A$ is cocartesian monoidal.   If this fiberwise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to any other cocartesian monoidal structure on the fibers, by the universal property of colimits. Therefore, we can simplify the above assumption to that in \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor, although we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$. For example, in that case, the coproducts induced in the Grothendieck category from the fiberwise ones as in \cref{eq:fibrewisecoprod} are equivalently given by \cref{eq:explicitstructure2}.

@@

Building up to an isomorphism of the above described double categories, we now define a double functor
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp
\end{displaymath}
which on vertical categories acts as the identity $\lE_0 = \id_{\A}$. Given a horizontal 1-cell in $_L\lCsp(\inta F)$ as in \cref{eq:scsphor1cell}, $\lE_1$ maps it to the decorated cospan $i\maps a\to m\leftarrow b:o$ with decoration $x\in F(m)$,
whereas a 2-morphism of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} is mapped to the underlying map of cospans in $\A$
\begin{displaymath}
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\end{displaymath}
in $\A$ along with the map $\tau\maps Fk(x)\to x'$ in $F(m')$ coming from the middle arrow above, namely an $F$-decorated cospan 2-morphism. 
With these definitions, $\lE$ is indeed a double functor as in \cref{def:doublefun}: both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets, and there exist globular 2-isomorphisms $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(N)\odot \mathbb{E}(M)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b,\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn)
\end{equation}
whereas first composing $M$ and $N$ in $_L\lCsp(\inta F)$ using pushouts in $\inta F$ as in \cref{eq:globalpushout} produces 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(u_m)x+_{(\bot_{m+_bn})}F(u_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $u$ are the canonical maps into the pushout.
Under $\lE$ this is mapped to the underlying $\A$-cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(u_m)x+F(u_n)y$, since a pushout over an initial object is really a coproduct. That decoration is isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}
F(m)\times F(n)\ar[rr,"\phi"]\ar[d,"F(u_m)\times F(u_n)"']\ar[dr,phantom,"\cong"] && F(m+n)\ar[d,"F(\psi)"]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(u_m+u_n)"description] \\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the bottom composite is the fiberwise coproduct in $F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand side isomorphism is due to pseudonaturality of $\phi$ whereas the right-hand side isomorphism follows from universal properties and pseudofunctoriality of $F$.
Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibers. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}. %Do we want to check that E is a double functor one last time?

@@

\begin{proof}
This follows by applying Shulman's result \cite[Theorem 1.2]{Shulman2010} to the fibrant symmetric monoidal double category $F\lCsp$.
\end{proof}

@@

Decorated cospans were introduced by Fong \cite{Fong,FongThesis} to describe open systems as cospans equipped with extra data.  They were then applied to open electrical circuits \cite{BF}, Markov processes \cite{BFP}, and chemical reaction networks \cite{BP}.  Unfortunately, some of these applications were marred by technical flaws, which were later fixed using structured cospans \cite[Section 5]{BC2}. Here we explain how to also fix them using our new decorated cospans, since they provide another another solution to these problems. Below we compare the two approaches in applications to graphs, electrical circuits, Petri nets, reaction networks and dynamical systems. In many cases, \cref{thm:equiv} shows that the structured and decorated cospan approaches are equivalent; that is why \cref{subsec:graphs,subsec:circuits,subsec:petri} are on the brief side. However, in \cref{subsec:petrirates} we elaborate and show that open dynamical systems can only be treated using decorated cospans.

@@

We can easily build a double category with `open graphs' as horizontal 1-cells using the machinery of structured cospans, see \cite[Section 5]{BC2}.  Let $L \maps \Fin\Set \to \Graph$ be the functor that assigns to a finite set $N$ the \define{discrete graph} on $N$: the graph with no edges and $N$ as its set of vertices. Both $\Fin\Set$ and $\Graph$ have finite colimits, and the functor $L \maps \Fin\Set \to \Graph$ is left adjoint to the forgetful functor $R \maps \Graph \to \Fin\Set$ that assigns to a graph $G$ its underlying set of vertices $R(G)$. Thus, using structured cospans and appealing to \cref{thm:SC}, we get a symmetric monoidal double category $_L \lCsp(\Graph)$ in which:
\begin{itemize}
\item objects are finite sets,
\item a vertical 1-morphism from $X$ to $Y$ is a function $f \maps X \to Y$,
\item a horizontal 1-cell from $X$ to $Y$ is an \define{open graph} from $X$ to $Y$, meaning a cospan in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a commuting diagram in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(Y_1)$};
\node (A') at (0,-1) {$L(X_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(Y_2)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
\end{itemize}
Here is an example of an open graph:
\[
\scalebox{0.8}{
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [contact] (n1) at (-2,0) {$\bullet$};
		\node [style = none] at (-2.1,0.3) {$n_1$};
		\node [contact] (n2) at (0,1) {$\bullet$};
		\node [style = none] at (0,1.3) {$n_2$};
		\node [contact] (n3) at (0,-1) {$\bullet$};
		\node [style = none] at (0,-1.3) {$n_3$};
		\node [contact] (n4) at (2,0) {$\bullet$};
		\node [style = none] at (2.1,0.3) {$n_4$};

@@

We can make $F$ into a symmetric lax monoidal pseudofunctor $F \maps (\Fin\Set, +) \to (\Cat,\times)$ by equipping it with suitable functors 
\[   \phi_{N,N'} \maps F(N) \times F(N') \to F(N+N'), \qquad \phi_0 \maps F(\emptyset) \to \one \in \Cat. \]  
{\chris Shouldn't the laxator for the units go in the other direction??}
